---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# smvr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/smvr)](https://CRAN.R-project.org/package=smvr)
<!-- badges: end -->

Simple implementation of [Semantic Versioning 2.0.0](https://semver.org/) on R.

smvr provides a lightweight, fully vectorized R class for Semantic Versioning.
It enables type-safe manipulation, comparison, and sorting of version numbers,
and integrates smoothly with tidyverse tools.

## Installation

Install the latest release version from CRAN:

```r
install.packages("smvr")
```

The development version can be installed from GitHub:

```r
pak::pak("eitsupi/smvr")
```

## Example

```{r example}
library(smvr)

# Parse version characters into smvr objects
v <- parse_semver(c("1.0.0", "1.0.0-alpha.2", "1.0.0-alpha.10", "1.0.1+20250621", "0.9.0"))

# Sort versions
sort(v)

# Can compare with string notation
v["1.0.0-alpha" < v & v < "1.0.0"]

# Works with tibble data frame and dplyr
tibble::tibble(version = v) |>
  dplyr::arrange(version) |>
  dplyr::mutate(
    `>= 1.0.0` = version >= "1.0.0",
    `pre-release` = is_pre_release(version),
  )
```

## Features

- Fully vectorized Semantic Versioning class.
- Type-safe comparison and sorting.
- Tidyverse compatibility (`{tibble}`, `{dplyr}`, etc.).
- No dependencies except `{vctrs}`.

## Related Works

- The numeric version class vector can be crated with `numeric_version()` in base R works well for versions
  that only have MAJOR.MINOR.PATCH. But it does not support pre-release identifiers of SemVer,
  so in the case of including pre-release versions, it is not suitable.
- The [semver](https://CRAN.R-project.org/package=semver) package is a wrapper for a C++ SemVer parser.
  The class provided by this package is a special list, which does not work well with `{tibble}` and `{dplyr}`.
- The [semverutils](https://CRAN.R-project.org/package=semverutils) package has a SemVer parser written in R.
  It creates a single version as an `{R6}` object, which is not vectorized.
  Also, at the moment (version 0.1.0, published 2020-02-22 on CRAN),
  it has a bug in comparing pre-release versions[^semverutils], so it does not work correctly.

[^semverutils]: <https://github.com/ajwtech/semverutils/issues/2>
