---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# smvr

<!-- badges: start -->
<!-- badges: end -->

Simple implementation of [Semantic Versioning 2.0.0](https://semver.org/) on R.

smvr provides a `{vctrs}`-based vector class `smvr` for semantic versions.
It enables type-safe manipulation, comparison, and sorting of semantic versions.

## Installation

You can install the development version of smvr with:

```r
pak::pak("eitsupi/smvr")
```

## Example

```{r example}
library(smvr)

# Parse version characters into smvr objects
v <- parse_smvr(c("1.0.0", "1.0.0-alpha", "1.0.0-alpha.1", "1.0.1+20250621", "0.9.0"))

# Sort versions
sort(v)

# Compare versions
smvr(1, 0, 0) < smvr(1, 0, 1)

# Works with tibble data frame and dplyr
tibble::tibble(version = v) |>
  dplyr::arrange(version) |>
  dplyr::mutate(`>= 1.0.0` = version >= smvr(1))
```

## Features

- Create version objects with `smvr()`
- Parse version characters with `parse_smvr()`
- Use comparison operators (<, >, ==), `sort`, `order`, etc. (`{vctrs}` compatible)

## Known Limitations

- The number of pre-release identifier fields is limited to 5.
- When comparing vectors, the type for pre-release identifiers is determined for the whole vector.
  This means that if any identifier is alphanumerical, all are compared as strings,
  so numeric comparison may not be used for some fields.

```{r limitations}
# Only 5 pre-release fields are supported:
try(parse_smvr("1.2.3-a.b.c.d.e.f"))

# Numeric comparison may not be used if any field is alphanumerical:
# For example, the third value "1.0.0-a.b" make the third pre-release identifier alphabetical,
# so all pre-release identifiers are compared as strings.
prerelease_include_alphabet <- parse_smvr(c("1.0.0-a.1", "1.0.0-a.2", "1.0.0-a.b"))
prerelease_include_alphabet[2]
# All are compared as strings, so "10" < "2" (alphabetical order)
prerelease_include_alphabet[2] < parse_smvr("1.0.0-a.10")

# If all pre-release identifiers are numeric, they can be compared numerically:
prerelease_only_numeric <- parse_smvr(c("1.0.0-a.1", "1.0.0-a.2", "1.0.0-a.3"))
prerelease_only_numeric[2]
prerelease_only_numeric[2] < parse_smvr("1.0.0-a.10")
```
